---
title: "xG Model v2"
author: "FT"
date: "15/08/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
Data acquired using the NHL scraper developed by Evolving Hockey
https://github.com/evolvingwild/evolving-hockey

Model built on the example of an xG model by Matthew Barlowe 
https://rstudio-pubs-static.s3.amazonaws.com/311470_f6e88d4842da46e9941cc6547405a051.html 
(NHL Expected Goals Model - Matthew Barlowe)

# Set up
## Packages
```{r setup, include=FALSE}
using<-function(...) {
  libs<-unlist(list(...))
  req<-unlist(lapply(libs,require,character.only=TRUE))
  need<-libs[req==FALSE]
  if(length(need)>0){ 
    install.packages(need)
    lapply(need,require,character.only=TRUE)
  }
}

using('ggplot2','dplyr','rstudioapi','data.table','ggthemes','tidyr','lubridate','scales','grid','pracma','mlr','mclust','mlbench','kernlab','ranger','cmaes','e1071','devtools','readr','stringr','foreach','jsonlite','rvest','xml2','RCurl','xgboost','fastDummies')

if (Sys.getenv("JAVA_HOME")!="")
  Sys.setenv(JAVA_HOME="")

memory.limit(size=56000)

options(digits=5)

```

## Directory
```{r}
current_path<-getActiveDocumentContext()$path
setwd(dirname(current_path ))
print(getwd())

```

# Data
## Loading in data
```{r}
data.18_19<-readRDS(paste0(getwd(),'/Data/2018-2019_PbP_Data.rds'))
data.19_20<-readRDS(paste0(getwd(),'/Data/2019-2020_PbP_Data.rds'))

```

## Separating the different dataframes
```{r}
# game_info_df_new<-pbp_scrape$game_info_df           ## game information data
# pbp_base_new<-pbp_scrape$pbp_base                   ## main play-by-play data
# pbp_extras_new<-pbp_scrape$pbp_extras               ## extra play-by-play data
# player_shifts_new<-pbp_scrape$player_shifts         ## full player shifts data
# player_periods_new<-pbp_scrape$player_periods       ## player TOI sums per period
# roster_df_new<-pbp_scrape$roster_df                 ## roster data
# scratches_df_new<-pbp_scrape$scratches_df           ## scratches data
# event_summary_df_new<-pbp_scrape$events_summary_df  ## event summary data
# scrape_report<-pbp_scrape$report                    ## scrape report

df.pbp_18_19<-data.18_19$pbp_base
df.pbp_19_20<-data.19_20$pbp_base

```

## Data exploration
```{r}
unique(df.pbp_18_19$event_type)
# Fenwick events are GOAL,MISS,SHOT

unique(df.pbp_18_19$game_strength_state)
# E means Extra, so Ev5 is 6v5, but can also be the start of the game before other team gets on ice

unique(df.pbp_18_19$event_detail)
# When its a fenwick event, this is where the shot type is recorded

```

# Building the dataset
## Adding basic flags and diff time
```{r}
df.pbp_18_19<-df.pbp_18_19%>%
  filter(game_period<5)%>%
  mutate(goal_flag = ifelse(event_type == 'GOAL', 1, 0),
         home_flag = ifelse(event_team == home_team, 1, 0),
         period_id = paste0(game_id,game_period))%>%
  group_by(period_id)%>%
  mutate(time_diff = game_seconds - lag(game_seconds))

df.pbp_18_19$time_diff[is.na(df.pbp_18_19$time_diff)]<-0

```

## Adding rebound and rush flag
```{r}
df.pbp_18_19<-df.pbp_18_19%>%
  mutate(rebound_flag = ifelse(time_diff<3
                               & event_type %in% c('SHOT','GOAL','MISS')
                               & lag(event_type) %in% c('SHOT','GOAL','MISS')
                               & event_team == lag(event_team), 1, 0),
         rush_flag = ifelse(time_diff<4
                            & lag(abs(coords_x))<25
                            & event_type %in% c('SHOT','GOAL','MISS'), 1, 0))%>%
  filter(event_type %in% c('SHOT','GOAL','MISS'))

df.pbp_18_19$rebound_flag[is.na(df.pbp_18_19$rebound_flag)]<-0
df.pbp_18_19$rush_flag[is.na(df.pbp_18_19$rush_flag)]<-0

```

## Adding shot location, angle, and distance
```{r}
df.pbp_18_19<-df.pbp_18_19%>%
  filter(coords_x != 'NA' & coords_y != 'NA')%>%
  mutate(coords_y = ifelse(coords_x < 0, -1*coords_y,coords_y),
         coords_x = abs(coords_x),
         shot_angle = (asin(abs(coords_y)/sqrt((87.95 - abs(coords_x))^2 + coords_y^2))*180)/3.14159,
         shot_angle = ifelse(abs(coords_x) > 88, 90 + (180-(90 + shot_angle)), shot_angle),
         shot_distance = sqrt((87.95 - abs(coords_x))^2 + coords_y^2))

```

## Looking at shot types closely
### Dealing with NA
```{r}
df.pbp_18_19%>%
  group_by(event_detail)%>%
  summarise(n())

df.pbp_18_19%>%
  filter(is.na(event_detail)==T)

# What are NA shot types? 7 goals out of 10 shots! Will be classified as "Other"

df.pbp_18_19$event_detail[is.na(df.pbp_18_19$event_detail)]<-'Other'
```

### Difference between Tip-In vs Deflected
```{r}
# I am a bit worried tip-in might be a deflected shot that was a goal
df.pbp_18_19%>%
  filter(event_detail %in% c('Tip-In','Deflected'))%>%
  group_by(event_detail)%>%
  summarize(goal_count = sum(goal_flag),
            goal_proportion = mean(goal_flag))

# No worries to have here. Maybe deflected is a pass that goes off a skate while tip in an intentionally deflected shot
```

## Removing empty net goals
```{r}
df.pbp_18_19<-df.pbp_18_19%>%
  mutate(empty_net_goal_flag = ifelse((event_type == 'GOAL'
                                       & home_flag == 1
                                       & is.na(away_goalie) == T)
                                      |(event_type == 'GOAL'
                                       & home_flag == 0
                                       & is.na(home_goalie) == T),1,0))%>%
  filter(empty_net_goal_flag != 1)

```

## Selecting the data to train the model on
```{r}
df.train<-df.pbp_18_19%>%
  select(period_id,
         event_index,
         event_detail,
         rebound_flag,
         rush_flag,
         coords_x,
         coords_y,
         shot_angle,
         shot_distance,
         goal_flag)%>%
  mutate(event_detail = as.factor(gsub('-','_',event_detail)),
         goal_flag=as.factor(goal_flag))

df.train<-dummy_cols(df.train, select_columns = 'event_detail')

df.train.id<-df.train%>%
  select(period_id,
         event_index)

df.train<-df.train[,-c(1,2,3)]
```


# Building the model
```{r}
hyper_grid<-makeTuneControlGrid()
kFold<-makeResampleDesc('RepCV', folds = 10, reps = 50)
task.xG<-makeClassifTask(data = df.train, target = 'goal_flag', positive = '1')
```

## XGboost
```{r}
# listLearners()$class

# getLearnerParamSet('classif.xgboost')
# xg_boost.hyper_param<-makeParamSet(makeNumericParam('eta',lower=0,upper=1),
#                                    makeDiscreteParam('booster',values=c('gbtree','dart')),
#                                    makeNumericParam('gamma',lower=0.1,upper=100),
#                                    makeIntegerParam('max_depth',lower=2,upper=15))

xg_boost.hyper_param<-makeParamSet(makeIntegerParam('max_depth',lower=15,upper=25))

xg_boost.learner<-makeLearner('classif.xgboost')

xg_boost.tune<-tuneParams('classif.xgboost', task = task.xG, resampling = kFold, par.set = xg_boost.hyper_param, control = hyper_grid, measures = list(tpr,f1))

# listMeasures('classif', properties = 'classif')
# xg_boost.CV<-resample(learner = xg_boost.learner, task = task.xG,resampling = kFold)

```


```{r}

```




